### 2022-09-04 Papers/ICCC-2023/sections/abstract.tex
stuff

CTFE first approach places the interpreter as the central component of the compiler.
Translation of user code into the executable form is done using an interpreted module, written in the target language, and uses the same data structures accessible to the programmer.
This boosts the flexibility and extensibility of the compiler and ensures the completness of the available metadata and the ability to execute any code at compile-time.

CTFE first pattern is designed with compiling low-level, non-garbage-collected, reflection-enabled languages in mind.
It was created in order to compile C-=-1, which aimed to allow the programmer to execute any code at compile time, as well as analyze and modify the program structure.
Thanks to the extensibility of the designed compiler, programs written in C-=-1 can also generate marshalling bindings for other languages and support a variety of unexpected programming paradigms.

>>>>

This paper proposes a new approach to compiler architecture that places the interpreter as the central component of the compiler.
Translation of user code into the executable form is done using an interpreted module, written in the target language, and uses the same data structures accessible to the programmer.
This boosts the flexibility and extensibility of the compiler and ensures the completness of the available metadata and the ability to execute any code at compile-time.

Compile Time Function Execution (CTFE) First pattern is designed with compiling low-level, non-garbage-collected, reflection-enabled languages in mind.
It was created in order to compile C-=-1, which aimed to allow the programmer to execute any code at compile time, as well as analyze and modify the program structure.
Thanks to the extensibility of the designed compiler, programs written in C-=-1 can also generate marshalling bindings for other languages and support a variety of unexpected programming paradigms.

### 2022-09-04 Papers/ICCC-2023/sections/implementation.tex
stuff

The first CTFE First was created for the C-=-1 language, using out-of-the-box parser generator and backend.

\subsection{Interpreter data structures}

Data structures of the C-=-1 Interpreter have been designed ease of development and debugging in mind.
They are thus not particularly efficent.

@@

The most prymitive types within the hierarchy are \lstinline{StringValue} and \lstinline{IntegerValue}.
They are simple wrappers for strings and integers.
Floating point numbers were not implemented as they were not nessecary for implementation of a basic compiler.

@@

There are multiple types of reference within the C-=-1 interpreter.
The most basic pointer type is a reference to C-=-1 value.

@@

>>>>

The first CTFE First was created for the C-=-1 language, using out-of-the-box parser generator and backend.
The most important aspect of implementing a CTFE First compiler is the design of the data structures, described in chapter \ref{data_structures}, that will be used by the interpreter.

Languages that seek to exploit CTFE First approach in design of their compilers, should define a set of data structures to describe user code, i.e. an Semantic Model.
It will allow the programmer to interact and mainpulate the structure of the program at compile-time.

\subsection{Interpreter data structures}
\label{data_structures}
Data structures of the C-=-1 Interpreter have been designed ease of development and debugging in mind.
They are thus not particularly efficent.

@@

The most prymitive types within the hierarchy are \lstinline{StringValue} and \lstinline{IntegerValue}.
They are simple wrappers for strings and integers, present in the host language.
Floating point numbers were not implemented as they were not nessecary for implementation of a basic compiler.

@@

There are multiple types of reference within the C-=-1 interpreter.
The most basic pointer type is a reference to C-=-1 value.
It was realised as a pointer to the owning pointer of the value.
The reasoning behind this decision was to allow assigning to the referenced value.

@@

A major motivation in creating CTFE First was the ability to support languages with compile-time metaprogramming.
This includes reflection and modification of the code being compiled.
User program had to be represented as a complete and modifiable object, using the interpreters data structures.

C-=-1 language model divides the user program into assemblies.
They represent an individual program package: a library or an executable file.
The compiler is invoked to compile an assembly, together with its dependencies.
Assembly is the root object of C-=-1 program model, it stores a list of assemblies it depends on and the root namespace of the package it represents.
The remainder of user code is organised around namespaces, types, functions and fields.
These parts of the model are repesented by native classes of the host language and form the basis for the rest of the model.

The most complex part of the model is the representation of the function body.

### 2022-09-04 Papers/ICCC-2023/sections/introduction.tex
stuff

Compile-time function execution is an aspect of a programming language, that allows the programmer to execute code at compile-time.
The goals and capabilities of CTFE depend on the language and is discussed in chapter \ref{related-work}, however it plays a secondary role in the compilation process.

CTFE first builds the compiler around the interpreter component.
The role of the compiler is to construct the semantic model of the program being compiled and pass it as data to the interpreted Compiler-Intreface module.
Compiler-Interface then tranforms the model into the backend's assembly language to be compiled to executable.
The design of the language that sparked this approach was described in chapter \ref{language-design}.
Chapter \ref{compiler-design} deals with structure of a CTFE First compiler and how its components interact.

>>>>

Compile-time function execution (CTFE) is an aspect of a programming language, that allows the programmer to execute code at compile-time.
It has been gaining popularity with low level programming langauges such as Rust or C++.
The goals and capabilities of CTFE depend on the language and are discussed in chapter \ref{related-work}, however it plays a secondary role in the compilation process.

This paper proposes a new approach to compiler construction that places Compile Time Function Execution capabilities as the first priority.
This architecture is called CTFE First.
CTFE First builds the compiler around the interpreter component.
The role of the compiler is to construct the semantic model of the program being compiled and pass it as data to the interpreted Compiler-Intreface module.
Compiler-Interface then tranforms the model into the backend's assembly language to be compiled to executable.
The design of the language that sparked this approach was described in chapter \ref{language-design}.
Chapter \ref{compiler-design} deals with structure of a CTFE First compiler and how its components interact.

### 2022-09-04 Papers/ICCC-2023/sections/realted-work.tex
Chapter layout
stuff

>>>>

Rust language compiler is the most similar in capability, to what was demonstrated with C-=-1, as a feature of CTFE First.
It allows the user to write code that performs some transformations of the program, and interact with the evnoirment during the build process.
The two relevant features are build scripts and macros.

Rust macros, mutch like the classic C-style preprocessor macros, generate additional code as text.
The major difference between the systems is that Rust macros operate on tokens rather than text and use syntax similar to regular Rust code.
This mechanism is powerfull, allowing the user to rewrite code to avoid repetition, simplify certain tasks and create new syntax.
They are however unable to reflect on the program as a whole or obtain mutch of the information available to the compiler.

Build scripts, on the other hand, are programs that execute prior to compilation of the main package.
They are meant to prepare the envoirment for building the program, by for example compiling its external dependencies.
They do not have any information about the structure of the program being compiled.
If a task, done by a build script, were to require such data, it would have to analyze the code by itself, without any help from the compiler.
This makes application such as automatically genertaing bindings for other programming languages significantly more difficult.
